Assessing the Replicability of Psychological Science Through Pedagogy

### 2016-08-22 paper/paper.Rmd
appendix

@@

@@

@@

@@

key statistic in original

@@

@@

@@

@@

@@

>>>>

This paper investigated motivations to save for retirement, and specifically whether changing participants' relationships to the future (making it feel more immanent) would cause them to save more. Our target was Study 4, which found that seeing the time to retirement in days (10,950) rather than years (30) caused participants to say they would begin saving sooner, even controlling for age, education, and income in a linear model (there was also a manipulation of whether saving was incremental that did not result in an effect). We failed to find a significant relationship between the time metric manipulation and saving time. However, when we did not include demographic controls, we did find a marginally-significant relationship. In exploratory analyses, we also observed an un-predicted effect of income such that participants who reported higher income selected to start saving sooner.

@@

The goal of this paper was to examine whether people are accurately calibrated in their estimates of how others will respond to their attempts at self-promotion. Our target was Experiment 3, which showed that participants, who were instructed to describe themself in such a way that others would like them, expected to be liked more -- but were actually liked less -- than participants who were asked to describe themselves but not additionally instructed to maximize others’ interest in meeting them. The authors performed a series of regressions on four dependent variables capturing judgements about profile writers (interest, liking, bragging, and success), analyzing these as a function of condition (control vs. "maximize interest") and who the evaluator was (the writer vs. an independent sample). Power analysis for this design was quite complex because the procedure had two stages: in a first stage participants generated descriptions and predicted judgments, while in a second stage a separate sample rated the descriptions.

We were faced with a choice: either we could have re-rated the descriptions gathered in the original study (by contacting the author and treating these as "materials" for the study), effectively replicating only stage two, or we could redo the entire two-step procedure. We elected to conduct a replication of the full procedure, but in our replication to decrease cost we focused on the liking judgments, which appeared to be central to the authors' conclusions. Our key test was the interaction of rater (writer vs. independent) and condition.^[We note that this project was submitted late and the student failed to preregister this analysis formally; however, the pre-written analysis protocol went through pre-data collection review by MCF.]  However, to power even this smaller study adequately was outside of our budget, so instead of having each judge rate a subset of 10 out of 100 total profiles, we collected a smaller sample of profiles (18) and had every profile rated by each judge. We failed to find the predicted interaction in this new sample. Instead, we found a main effect such that stage one participants predicted that raters would like them more than the stage two participants actually did. Despite our relatively high statistical power in the main test, however, we may have suffered from low power at the item level due to the relatively small number of profiles we collected in the replication study.

@@

This paper tested the theory that persistent object representations could assist in spatial navigation, using a navigation paradigm in which participants used keypresses to move though a grid of pictures that were visible only one at a time. In Experiment 1, our target, found that participants located targets faster when navigation involved persistence cues (via sliding animations) than when persistence was disrupted (via temporally matched fading animations). The key test for our replication was a simple *t*-test comparing speed to find the target in the two conditions (slide vs. fade), not controlling for learning across the duration of the experiment. We failed to find this overall difference. However, in an exploratory followup analysis using a linear mixed effect model, we did find a highly-significant effect of condition when controlling for learning epoch. We speculate that the very large effect found by the original authors may have led them to select a simpler analysis that did not control for learning effects. Our data -- which were more variable -- required this control.

@@

This paper reported that the relationship between math anxiety and math performance follows an inverted-U pattern among college undergraduate students with higher intrinsic math motivation, whereas this same relationship is linearly negative among students with lower intrinsic math motivation. We targeted Study 2, the authors' own replication study, in which participants filled out three surveys about their math motivation and anxiet and completed a math task to measure performance. The key effect we targeted was the interaction of math anxiety squared and math motivation in predicting performance. We failed to find evidence for this effect. In contrast to the original study, we found a simple linear trend such that performance decreased with increasing math anxiety for both high and low math-motivation groups.

@@

This paper tested the hypothesis that face typicality affects trustworthiness judgements such that the most typical face will be judged to be the most trustworth. In Study 1, our target, faces of varying typicality were rated on their trustworthiness and attractiveness; while trustworthiness peaked at the most typical face, less typical faces were judged more attractive. While the original class projects selected a global, item-wise regression model as the key statistical test of interest, the instructor team (post-data collection) determined that the fit of this model did not in fact correspond most closely with the authors' stated hypothesis. We thus selected the interaction between face typicality and rating condition as our key test statistic and used the more standard by-participant regression model (which the original paper also reported subsequent to the by-items model) as the target model. Our replication study successfully showed this same interaction. We note that, despite the significant interaction, in our study trustworthiness judgments did not change substantially with typicality (though attractiveness did) and trustworthiness did not peak numerically at the most typical face. Thus, despite replicating one key statistical test, our results are at odds with some of the original papers' claims.

@@

This paper investigated vocal cues to social hierarchy. Our target for replication was Experiment 2, which assessed whether participants used particular acoustic cues to make inferences about speakers' level of social hierarchy, particularly that participants from the high-rank condition in Experiment 1 were rated as more highly ranking. This paper employed a two-stage procedure in which participant-generated materials were rated by an independent sample. However, in contrast to the study of Scopelliti, Lowenstein & Vosgerau (2015), because materials were collected in Experiment 1 and were available openly, we elected to replicate only Experiment 2, the judgment study. The key test statistic was the main effect of hierarchy condition on "behavior score" (an index of whether speakers would engage in high-status behaviors). The authors were also interested in which particular vocal cues predicted participants judgments, but we judged these analyses to be more descriptive and exploratory. We successfully replicated the main effect of condition on hierarchy judgment, and additionally found a comparable main effect of speaker gender.

@@

This paper investigated how perceived knowledge affects "overclaiming" (claiming knowledge that you do not have. We targeted Study 1b, which asked participants to rate their knowledge about finances, and then asked them to rate their knowledge about a set of financial terms, some of which were invented (and hence for which a positive knowledge statement would necessarily be an overclaim). The original study found that participants with higher self-rated financial knowledge also overclaimed at a higher rate. The key test statistic for this study was the coefficient in a regression model predicting overclaiming as a function of claimed personal financial knowledge, controlling for accuracy in the true finacial knowledge questions. We successfully replicated this effect, despite our use of a much smaller sample (due to the extremely high post-hoc power of the original study).

@@

This paper examined whether judgments of creativity are influenced by the gender of the creator. In Study 1, they tested the hypothesis that "out of the box" creativity was more associated with masculine traits than feminine traits by manipulating the definition of creativity that participants saw (convergent vs. divergent) and asking participants to judge the centrality of different traits to that definition. The original pre-registration selected a subsidiary analysis of a main effect as the key test statistic, but in post-data collection review the instructor team decided that a test of the interaction between trait gender and definition condition was closer to the central theoretical claim of the paper; we use this latter test as the key test statistic. Our replication study found an significant interaction of similar magnitude.

@@

This paper investigated whether encouraging participants to think about their personal legacy could increase concern for the environment. The main study of the paper investigated whether a legacy priming writing exercise would increase a variety of environmental measures, including donations to an environmental charity, proenvironmental intentions, and climate change beliefs. The key test we selected was the effect of condition (priming vs. control) on behavioral intentions in an analysis of variance. We observed a marginally significant effect in the same direction in our replication, as well as support for the mediating relationship of legacy motives on beahvioral intentions. Our study may have been underpowered due to budgetary constraints; we ran a sample of comparable size to the original sample of 312 participants. We also may have observed a smaller effect size due to the relatively smaller amount of time our participants spent on the legacy prime writing exercise.

@@

This paper explored constraints on visual working memory in the context of mental rotation. In Experiment 1a, participants perfomed a verbal suppression task to control for use of verbal encoding while attempting to remember a cross with four colored bars, which sometimes switched colors; performing mental rotation significantly impaired ability to detect these bar swaps. To port the task to an online framework we decreased the number of trials and increased the number of participants we tested. The key test statistic was a comparison of K (a metric of the number of visual features remembered) between the rotation and no-rotation conditions. We observed a marginally-significant effect in our replication study. Our study showed a floor effect on capacity such that participants remembered on average very little; it is possible that small details in the displays (or the differing sample available online) led to this floor effect.
